---
title: "HS project"
subtitle: "Zoom in IBL and mORS cells"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This file is used to analyse the immune cells dataset.

```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(ComplexHeatmap)
library(org.Mm.eg.db)

.libPaths()
```

# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "iblmors"
out_dir = "."
```

We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/", save_name, "_sobj.rds"))
sobj
```

We load the sample information :

```{r custom_palette_sample, fig.width = 6, fig.height = 6}
sample_info = readRDS(paste0(out_dir, "/../../1_metadata/hs_hd_sample_info.rds"))
project_names_oi = sample_info$project_name

graphics::pie(rep(1, nrow(sample_info)),
              col = sample_info$color,
              labels = sample_info$project_name)
```

Here are custom colors for each cell type :

```{r color_markers, fig.width = 10, fig.height = 1, class.source = "fold-hide"}
color_markers = readRDS(paste0(out_dir, "/../../1_metadata/hs_hd_color_markers.rds"))

data.frame(cell_type = names(color_markers),
           color = unlist(color_markers)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank(),
                 axis.text.x = element_text(angle = 30, hjust = 1))
```

This is the projection of interest :

```{r name2D}
name2D = "harmony_20_tsne"
```


# Visualization

## Gene expression

We visualize gene expression for some markers :

```{r plot_list_features, fig.width = 12, fig.height = 4}
features = c("percent.mt", "percent.rb", "nFeature_RNA")

plot_list = lapply(features, FUN = function(one_gene) {
  Seurat::FeaturePlot(sobj, features = one_gene,
                      reduction = name2D) +
    ggplot2::theme(aspect.ratio = 1) +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
    Seurat::NoAxes()
})

patchwork::wrap_plots(plot_list, ncol = 3)
```

## Clusters

We visualize clusters :

```{r see_clustering, fig.width = 6, fig.height = 4}
cluster_plot = Seurat::DimPlot(sobj, reduction = name2D, label = TRUE) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
cluster_plot
```

## Cell type

We visualize cell type split by sample :

```{r plot_split_dimred, fig.width = 12, fig.height = 7}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "project_name",
                                        group_by = "cell_type",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = color_markers,
                                        bg_pt_size = 0.5, main_pt_size = 0.5)

plot_list[[length(plot_list) + 1]] = cluster_plot

patchwork::wrap_plots(plot_list, ncol = 4) &
  Seurat::NoLegend()
```

## Cluster type

We summarize major cell type by cluster :

```{r cell_type_clusters}
cell_type_clusters = sobj@meta.data[, c("cell_type", "seurat_clusters")] %>%
  table() %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
cell_type_clusters = setNames(levels(sobj$cell_type)[cell_type_clusters],
                              nm = names(cell_type_clusters))

```

We define cluster type :

```{r table_cluster_type}
sobj$cluster_type = cell_type_clusters[sobj$seurat_clusters] %>%
  as.factor()
table(sobj$cluster_type, sobj$cell_type)
```

We subset `color_markers` :

```{r subset_colors}
color_markers = color_markers[levels(sobj$cluster_type)]
```


We compare cluster annotation and cell type annotation :

```{r see_cluster_type, fig.width = 10, fig.height = 5}
p1 = Seurat::DimPlot(sobj, group.by = "cell_type",
                     reduction = name2D, cols = color_markers) +
  ggplot2::labs(title = "Cell type") +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

p2 = Seurat::DimPlot(sobj, group.by = "cluster_type",
                     reduction = name2D, cols = color_markers) +
  ggplot2::labs(title = "Cluster type") +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

patchwork::wrap_plots(p1, p2, guides = "collect")
```

## Barplot

We make a barplot to compare IBL and ORS populations in HS vs HD samples. The proportion of IBL is indicated on top of bars.

```{r barplot_ibl_ors, fig.width = 7, fig.height = 5}
quantif = table(sobj$sample_identifier,
                sobj$cluster_type) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Sample", "cell_type", "nb_cells")) %>%
  dplyr::group_by(Sample) %>%
  dplyr::mutate(total_cells = sum(nb_cells)) %>%
  as.data.frame() %>%
  dplyr::filter(cell_type == "IBL") %>%
  dplyr::mutate(prop_ibl = nb_cells / total_cells) %>%
  dplyr::mutate(prop_ibl = 100*round(prop_ibl, 4))

sobj$seurat_clusters = factor(sobj$seurat_clusters,
                              levels = names(sort(cell_type_clusters)))

aquarius::plot_barplot(df = table(sobj$sample_identifier,
                                  sobj$seurat_clusters) %>%
                         as.data.frame.table() %>%
                         `colnames<-`(c("sample_identifier", "clusters", "nb_cells")),
                       x = "sample_identifier", y = "nb_cells", fill = "clusters",
                       position = position_fill()) +
  ggplot2::scale_fill_manual(values = c(colorRampPalette(c("chartreuse1", "chartreuse4"))(table(sort(cell_type_clusters))["IBL"]),
                                        colorRampPalette(c("cadetblue1", "cadetblue4"))(table(sort(cell_type_clusters))["ORS"])),
                             breaks = names(sort(cell_type_clusters)),
                             name = "Cell type") +
  ggplot2::geom_label(data = quantif, inherit.aes = FALSE,
                      aes(x = .data$Sample, y = 1.05, label = .data$prop_ibl),
                      label.size = 0, size = 5)
```

# Differential expression

In this section, we perform DE between :

- inner bulge layer (IBL) and outer root sheath (ORS) populations
- healthy donors (HD) and HS patients (HS) within ORS population
- healthy donors (HD) and HS patients (HS) within IBL population
- cluster 5 vs rest of ORS, because this cluster is specific to HS compared to HD

We save the results in a list :

```{r list_results}
list_results = list()
```


## IBL vs ORS

We change cell identities to cluster type :

```{r ident_cluster_type}
Seurat::Idents(sobj) = sobj$cluster_type

table(Seurat::Idents(sobj), sobj$sample_type)
```

We identify specific markers for each population :

```{r de_clust0, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj, ident.1 = "IBL", ident.2 = "ORS")

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[["IBL_vs_ORS"]] = mark

dim(mark)
head(mark, n = 20)
```

There are `r nrow(mark)` genes differentially expressed. How many for each population ?

```{r makr_by_pop}
# IBL
mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  nrow()

# ORS
mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  nrow()
```

We represent the information on a figure :

```{r plot_de_pop, fig.width = 10, fig.height = 10}
mark$gene_name = rownames(mark)
mark_to_label = rbind(
  # up-regulated in IBL
  mark %>% dplyr::top_n(., n = 20, wt = avg_logFC),
  # up-regulated in ORS
  mark %>% dplyr::top_n(., n = 20, wt = -avg_logFC),
  # representative and selective for IBL
  mark %>% dplyr::top_n(., n = 20, wt = (pct.1 - pct.2)),
  # representative and selective for ORS
  mark %>% dplyr::top_n(., n = 20, wt = -(pct.1 - pct.2))) %>%
  dplyr::distinct()

ggplot2::ggplot(mark, aes(x = pct.1, y = pct.2, col = avg_logFC)) +
  ggplot2::geom_abline(slope = 1, intercept = 0, lty = 2) +
  ggplot2::geom_point() +
  ggrepel::geom_label_repel(data = mark_to_label,
                            aes(x = pct.1, y = pct.2, label = gene_name),
                            col = "black", fill = NA, size = 3, label.size = NA) +
  ggplot2::labs(title = "Differentially expressed genes",
                subtitle = "between IBL and ORS") +
  ggplot2::scale_color_gradient2(low = aquarius::color_cnv[1],
                                 mid = aquarius::color_cnv[2],
                                 high = aquarius::color_cnv[3],
                                 midpoint = 0) +
  ggplot2::theme_classic() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5),
                 plot.subtitle = element_text(hjust = 0.5))
```

We represent the best genes on a violin plot, group by cluster type and split by sample type :

```{r violin_de_pop, fig.width = 15, fig.height = 39}
mark_to_label = mark_to_label %>%
  dplyr::arrange(pct.1 - pct.2)

plot_list = lapply(mark_to_label$gene_name, FUN = function(gene) {
  p = Seurat::VlnPlot(sobj, features = gene, pt.size = 0.001,
                      group.by = "cluster_type",
                      split.by = "sample_type") +
    ggplot2::scale_fill_manual(breaks = c("HS", "HD"),
                               values = c("#C55F40", "#2C78E6")) +
    ggplot2::theme(axis.title.x = element_blank(),
                   axis.title.y = element_blank(),
                   legend.position = "none")
  return(p)
})

patchwork::wrap_plots(plot_list, ncol = 5)
```


On the figure, we can see that some specific markers for a population are indeed specific to a sample type rather than the whole population.

## Cluster 5 within ORS

We subset the Seurat object and change cell identities to sample type.

```{r sub_cluster5}
subsobj = subset(sobj, cluster_type == "ORS")
subsobj$seurat_clusters = base::droplevels(subsobj$seurat_clusters)
Seurat::Idents(subsobj) = subsobj$seurat_clusters

table(subsobj$seurat_clusters)
```

We identify specific markers for each population :

```{r de_clust5, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(subsobj, ident.1 = 5)

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[["cluster5_vs_ORS"]] = mark

dim(mark)
head(mark, n = 20)
```

There are `r nrow(mark)` genes differentially expressed. How many for each population ?

```{r mark_cluster5}
# Cluster 5
mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  nrow()

# Other ORS
mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  nrow()
```

We represent the information on a figure :

```{r plot_de_cluster5, fig.width = 10, fig.height = 10}
mark$gene_name = rownames(mark)
mark_cluster5 = rbind(
  # up-regulated in cluster 5
  mark %>% dplyr::top_n(., n = 20, wt = avg_logFC),
  # up-regulated in other ORS
  mark %>% dplyr::top_n(., n = 20, wt = -avg_logFC),
  # representative and selective for cluster 5
  mark %>% dplyr::top_n(., n = 20, wt = (pct.1 - pct.2)),
  # representative and selective for other ORS
  mark %>% dplyr::top_n(., n = 20, wt = -(pct.1 - pct.2))) %>%
  dplyr::distinct()

ggplot2::ggplot(mark, aes(x = pct.1, y = pct.2, col = avg_logFC)) +
  ggplot2::geom_abline(slope = 1, intercept = 0, lty = 2) +
  ggplot2::geom_point() +
  ggrepel::geom_label_repel(data = mark_cluster5,
                            aes(x = pct.1, y = pct.2, label = gene_name),
                            col = "black", fill = NA, size = 3, label.size = NA) +
  ggplot2::labs(title = "Differentially expressed genes",
                subtitle = "between cluster 5 and other ORS") +
  ggplot2::scale_color_gradient2(low = aquarius::color_cnv[1],
                                 mid = aquarius::color_cnv[2],
                                 high = aquarius::color_cnv[3],
                                 midpoint = 0) +
  ggplot2::theme_classic() +
  ggplot2::theme(plot.title = element_text(hjust = 0.5),
                 plot.subtitle = element_text(hjust = 0.5))
```

We represent a subset of genes on a heatmap :

```{r prep_genes_cluster5}
features_oi = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(abs(avg_logFC) > 0.5) %>%
  rownames()

length(features_oi)
```

We prepare the scaled expression matrix :

```{r mat_cluster5}
mat_expression = Seurat::GetAssayData(subsobj, assay = "RNA", slot = "data")[features_oi, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
mat_expression = as.matrix(mat_expression) # not sparse

dim(mat_expression)
```

We prepare the heatmap annotation :

```{r ha_top_cluster5}
ha_top = ComplexHeatmap::HeatmapAnnotation(
  cluster_type = subsobj$cluster_type,
  sample_type = subsobj$sample_type,
  cluster = subsobj$seurat_clusters,
  col = list(cluster_type = color_markers,
             sample_type = setNames(nm = c("HS", "HD"),
                                    c("#C55F40", "#2C78E6")),
             cluster = setNames(nm = levels(subsobj$seurat_clusters),
                                aquarius::gg_color_hue(length(levels(subsobj$seurat_clusters))))))
```

And the heatmap :

```{r heatmap_cluster5, fig.width = 15, fig.height = 30}
ht = ComplexHeatmap::Heatmap(mat_expression,
                             col = aquarius::color_cnv,
                             # Annotation
                             top_annotation = ha_top,
                             # Grouping
                             column_title = NULL,
                             cluster_rows = TRUE,
                             cluster_columns = TRUE,
                             show_column_names = FALSE,
                             # Visual aspect
                             show_heatmap_legend = TRUE,
                             border = TRUE)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```


## IBL : HS vs HD

We subset the Seurat object and change cell identities to sample type.

```{r sub_IBL}
subsobj = subset(sobj, cluster_type == "IBL")
Seurat::Idents(subsobj) = subsobj$sample_type

table(subsobj$sample_type)
```

We identify DE genes between HS and HD :

```{r de_ibl, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(subsobj, ident.1 = "HS", ident.2 = "HD")

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[["IBL_HS_vs_HD"]] = mark

dim(mark)
head(mark, n = 20)
```

## ORS : HS vs HD

We subset the Seurat object and change cell identities to sample type.

```{r sub_ORS}
subsobj = subset(sobj, cluster_type == "ORS")
Seurat::Idents(subsobj) = subsobj$sample_type

table(subsobj$sample_type)
```

We identify DE genes between HS and HD :

```{r de_ors, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(subsobj, ident.1 = "HS", ident.2 = "HD")

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[["ORS_HS_vs_HD"]] = mark

dim(mark)
head(mark, n = 20)
```


# Heatmap

We represent differentially expressed genes. First, we extract all DE genes :

```{r prep_genes}
features_oi = c(mark_to_label$gene_name,
                rownames(list_results$IBL_HS_vs_HD),
                rownames(list_results$ORS_HS_vs_HD)) %>%
  unique()

length(features_oi)
```

## Heatmap

We prepare the scaled expression matrix :

```{r mat}
mat_expression = Seurat::GetAssayData(sobj, assay = "RNA", slot = "data")[features_oi, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
mat_expression = as.matrix(mat_expression) # not sparse

dim(mat_expression)
```

We prepare the heatmap annotation :

```{r ha_top}
ha_top = ComplexHeatmap::HeatmapAnnotation(
  cluster_type = sobj$cluster_type,
  sample_type = sobj$sample_type,
  cluster = sobj$seurat_clusters,
  col = list(cluster_type = color_markers,
             sample_type = setNames(nm = c("HS", "HD"),
                                    c("#C55F40", "#2C78E6")),
             cluster = setNames(nm = levels(sobj$seurat_clusters),
                                aquarius::gg_color_hue(length(levels(sobj$seurat_clusters))))))
```

And the heatmap :

```{r heatmap, fig.width = 15, fig.height = 45}
sobj$cell_group = paste0(sobj$cluster_type, sobj$sample_type) %>%
  as.factor()

ht = ComplexHeatmap::Heatmap(mat_expression,
                             col = aquarius::color_cnv,
                             # Annotation
                             top_annotation = ha_top,
                             # Grouping
                             column_order = sobj@meta.data %>%
                               dplyr::arrange(cluster_type, sample_type, seurat_clusters) %>%
                               rownames(),
                             column_split = sobj$cell_group,
                             column_gap = unit(c(0.01, 2, 0.01), "mm"),
                             column_title = NULL,
                             cluster_rows = TRUE,
                             cluster_columns = FALSE,
                             show_column_names = FALSE,
                             # Visual aspect
                             show_heatmap_legend = TRUE,
                             border = TRUE)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```

# Save

We save the list of results :

```{r save_list_results}
saveRDS(list_results, file = paste0(out_dir, "/", save_name, "_list_results.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

