---
title: "OEP002321 dataset"
subtitle: "Combined dataset"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <<- Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This file is used to generate a dataset containing all individual datasets, without melanocytes.

```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(ComplexHeatmap)

.libPaths()
```


# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "hs_hd"
out_dir = "."
n_threads = 5 # for tSNE
```


We load the sample information :

```{r custom_palette_sample, fig.width = 5, fig.height = 5}
sample_info = readRDS(paste0(out_dir, "/../1_metadata/hs_hd_sample_info.rds"))
project_names_oi = sample_info$project_name

graphics::pie(rep(1, nrow(sample_info)),
              col = sample_info$color,
              labels = sample_info$project_name)
```

Here are custom colors for each cell type :

```{r color_markers, fig.width = 10, fig.height = 1, class.source = "fold-hide"}
color_markers = readRDS(paste0(out_dir, "/../1_metadata/hs_hd_color_markers.rds"))

data.frame(cell_type = names(color_markers),
           color = unlist(color_markers)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank())
```

We load the markers and specific colors for each cell type :

```{r cell_markers}
cell_markers = readRDS(paste0(out_dir, "/../1_metadata/hs_hd_cell_markers.rds"))
lengths(cell_markers)
```

We load markers to display on the dotplot :

```{r dotplot_markers}
dotplot_markers = readRDS(paste0(out_dir, "/../1_metadata/hs_hd_dotplot_markers.rds"))
dotplot_markers
```

# Make `r save_name` dataset

## Individual datasets

For each sample, we :

* load individual dataset
* look at cell annotation

We load individual datasets :

```{r sobj_list}
sobj_list = lapply(project_names_oi, FUN = function(one_project_name) {
  subsobj = readRDS(paste0(out_dir, "/../2_individual/datasets/",
                           one_project_name, "_sobj_filtered.rds"))
  return(subsobj)
})
names(sobj_list) = project_names_oi

lapply(sobj_list, FUN = dim) %>%
  do.call(rbind, .) %>%
  rbind(., colSums(.))
```


We represent cells in the tSNE :

```{r name2D}
name2D = "RNA_pca_20_tsne"
```


We look at cell type annotation for each dataset :

```{r cell_type_proj, fig.width = 14, fig.height = 8}
plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  mytitle = as.character(unique(one_sobj$project_name))
  mysubtitle = ncol(one_sobj)
  
  p = Seurat::DimPlot(one_sobj, group.by = "cell_type",
                      reduction = name2D) +
    ggplot2::scale_color_manual(values = color_markers,
                                breaks = names(color_markers),
                                name = "Cell Type") +
    ggplot2::labs(title = mytitle,
                  subtitle = paste0(mysubtitle, " cells")) +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5)) +
    Seurat::NoAxes()
  
  return(p)
})

plot_list[[length(plot_list) + 1]] = patchwork::guide_area()

patchwork::wrap_plots(plot_list, ncol = 4) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```


and clustering :


```{r clustering_proj, fig.width = 14, fig.height = 8}
plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  mytitle = as.character(unique(one_sobj$project_name))
  mysubtitle = ncol(one_sobj)
  
  p = Seurat::DimPlot(one_sobj, group.by = "seurat_clusters",
                      reduction = name2D, label = TRUE) +
    ggplot2::labs(title = mytitle,
                  subtitle = paste0(mysubtitle, " cells")) +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5)) +
    Seurat::NoAxes() + Seurat::NoLegend()
  
  return(p)
})

patchwork::wrap_plots(plot_list, ncol = 4)
```

## Melanocytes remomal

For each individual dataset, we remove melanocytes. First, we smooth cell type annotation at a cluster level :

```{r smooth_annotation}
sobj_list = lapply(sobj_list, FUN = function(one_sobj) {
  cluster_type = table(one_sobj$cell_type, one_sobj$seurat_clusters) %>%
    prop.table(., margin = 2) %>%
    apply(., 2, which.max)
  cluster_type = setNames(nm = names(cluster_type),
                          levels(one_sobj$cell_type)[cluster_type])
  
  one_sobj$cluster_type = cluster_type[one_sobj$seurat_clusters]
  
  ## Output
  return(one_sobj)
})
```

To locate melanocytes, we look at their score, cell type annotation, and clustering.

```{r plot_cell_type, fig.width = 12, fig.height = 27}
plot_list = lapply(sobj_list, FUN = function(one_sobj) {
  project_name = as.character(unique(one_sobj$project_name))
  plot_sublist = list()
  
  # Score
  plot_sublist[[1]] = Seurat::FeaturePlot(one_sobj, reduction = name2D,
                                          features = "score_melanocytes") +
    ggplot2::labs(title = project_name,
                  subtitle = "Melanocytes score") +
    Seurat::NoAxes() +
    ggplot2::scale_color_gradientn(colors = aquarius:::color_gene) +
    ggplot2::theme(aspect.ratio = 1,
                   plot.subtitle = element_text(hjust = 0.5))
  
  # Cell type
  plot_sublist[[2]] = Seurat::DimPlot(one_sobj,
                                      reduction = name2D,
                                      group.by = "cell_type",
                                      order = "melanocytes") +
    ggplot2::scale_color_manual(values = c("purple", rep("gray92", length(color_markers) - 1)),
                                breaks = c("melanocytes", setdiff(names(color_markers), "melanocytes"))) +
    ggplot2::labs(title = "Cell type annotation",
                  subtitle = paste0(sum(one_sobj$cell_type == "melanocytes"),
                                    " melanocytes")) +
    Seurat::NoAxes() + Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5))
  
  # Clusters
  plot_sublist[[3]] = Seurat::DimPlot(one_sobj,
                                      reduction = name2D,
                                      group.by = "seurat_clusters",
                                      label = TRUE) +
    ggplot2::labs(title = "Clusters") +
    Seurat::NoAxes() + Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5))
  
  # Cluster type
  plot_sublist[[4]] = Seurat::DimPlot(one_sobj,
                                      reduction = name2D,
                                      group.by = "cluster_type") +
    ggplot2::scale_color_manual(values = c("purple", rep("gray92", length(color_markers) - 1)),
                                breaks = c("melanocytes", setdiff(names(color_markers), "melanocytes"))) +
    ggplot2::labs(title = "Cluster annotation",
                  subtitle = paste0(sum(one_sobj$cluster_type == "melanocytes"),
                                    " melanocytes")) +
    Seurat::NoAxes() + Seurat::NoLegend() +
    ggplot2::theme(aspect.ratio = 1,
                   plot.title = element_text(hjust = 0.5),
                   plot.subtitle = element_text(hjust = 0.5))
  
  return(plot_sublist)
}) %>% unlist(., recursive = FALSE)

patchwork::wrap_plots(plot_list, ncol = 4)
```

We remove melanocytes based on cluster annotation :

```{r remove_melanocytes}
sobj_list = lapply(sobj_list, FUN = function(one_sobj) {
  one_sobj$is_of_interest = (one_sobj$cluster_type != "melanocytes")
  
  if (sum(one_sobj$is_of_interest) > 0) {
    one_sobj = subset(one_sobj, is_of_interest == TRUE)
  } else {
    one_sobj = NA
  }
  
  one_sobj$is_of_interest = NULL
  return(one_sobj)
})

lapply(sobj_list, FUN = dim) %>%
  do.call(rbind, .) %>%
  rbind(., colSums(.))
```

## Re-annotation

We remove melanocytes from annotation :

```{r remove_from_annot}
cell_markers = cell_markers[names(cell_markers) != "melanocytes"]
color_markers = color_markers[names(color_markers) != "melanocytes"]
dotplot_markers = dotplot_markers[names(dotplot_markers) != "melanocytes"]
```

We re-annote cells for cell type, since melanocytes have been removed :

```{r re_annot}
sobj_list = lapply(sobj_list, FUN = function(one_sobj) {
  # Remove old annotation
  one_sobj@meta.data[, grep(colnames(one_sobj@meta.data), pattern = "score", value = TRUE)] = NULL
  
  # Re-annot
  one_sobj = aquarius::cell_annot_custom(one_sobj,
                                         newname = "cell_type",
                                         markers = cell_markers,
                                         use_negative = TRUE,
                                         add_score = FALSE,
                                         verbose = TRUE)
  
  # Set factor levels
  one_sobj$cell_type = factor(one_sobj$cell_type, levels = names(cell_markers))
  
  return(one_sobj)
})
```


## Combined dataset

We combine all datasets :

```{r merge_datasets}
sobj = base::merge(sobj_list[[1]],
                   y = sobj_list[c(2:length(sobj_list))],
                   add.cell.ids = names(sobj_list))
sobj
```

We add again the correspondence between gene names and gene ID. Since all datasets have been aligned using the same transcriptome, we take the correspondence from one individual dataset.

```{r add_metafeatures}
sobj@assays$RNA@meta.features = sobj_list[[1]]@assays$RNA@meta.features[, c("Ensembl_ID", "gene_name")]

head(sobj@assays$RNA@meta.features)
```

We remove the list of objects :

```{r clean_sobj_list}
rm(sobj_list)
```

We keep a subset of meta.data and reset levels :

```{r sobj_set_factor_levels}
sobj@meta.data = sobj@meta.data[, c("orig.ident", "nCount_RNA", "nFeature_RNA", "log_nCount_RNA",
                                    "project_name", "sample_identifier", "sample_type",
                                    "Seurat.Phase", "cyclone.Phase", "percent.mt", "percent.rb",
                                    "cell_type")]

sobj$orig.ident = factor(sobj$orig.ident, levels = levels(sample_info$project_name))
sobj$project_name = factor(sobj$project_name, levels = levels(sample_info$project_name))
sobj$sample_identifier = factor(sobj$sample_identifier, levels = levels(sample_info$sample_identifier))
sobj$sample_type = factor(sobj$sample_type, levels = levels(sample_info$sample_type))
sobj$cell_type = factor(sobj$cell_type, levels = names(color_markers))

summary(sobj@meta.data)
```

# Processing

We remove genes that are expressed in less than 5 cells :

```{r filter_genes}
sobj = aquarius::filter_features(sobj, min_cells = 5)
sobj
```

## Metadata

How many cells by sample ?

```{r table_orig_ident}
table(sobj$project_name)
```

We represent this information as a barplot :

```{r barplot_count, fig.width = 10, fig.height = 5}
aquarius::plot_barplot(df = table(sobj$project_name,
                                  sobj$cell_type) %>%
                         as.data.frame.table() %>%
                         `colnames<-`(c("Sample", "Cell Type", "Number")),
                       x = "Sample", y = "Number", fill = "Cell Type",
                       position = position_fill()) +
  ggplot2::scale_fill_manual(values = unlist(color_markers),
                             breaks = names(color_markers),
                             name = "Cell Type")
```

This is the same barplot with another position :

```{r barplot_stack, fig.width = 10, fig.height = 5}
aquarius::plot_barplot(df = table(sobj$project_name,
                                  sobj$cell_type) %>%
                         as.data.frame.table() %>%
                         `colnames<-`(c("Sample", "Cell Type", "Number")),
                       x = "Sample", y = "Number", fill = "Cell Type",
                       position = position_stack()) +
  ggplot2::scale_fill_manual(values = unlist(color_markers),
                             breaks = names(color_markers),
                             name = "Cell Type")
```

## Projection

We normalize the count matrix for remaining cells and select highly variable features :

```{r normalization}
sobj = Seurat::NormalizeData(sobj,
                             normalization.method = "LogNormalize")
sobj = Seurat::FindVariableFeatures(sobj, nfeatures = 2000)
sobj = Seurat::ScaleData(sobj)

sobj
```

We perform a PCA :

```{r pca}
sobj = Seurat::RunPCA(sobj,
                      assay = "RNA",
                      reduction.name = "RNA_pca",
                      npcs = 100,
                      seed.use = 1337L)
sobj
```

We choose the number of dimensions such that they summarize 60 % of the variability :

```{r ndims}
stdev = sobj@reductions[["RNA_pca"]]@stdev
stdev_prop = cumsum(stdev)/sum(stdev)
ndims = which(stdev_prop > 0.60)[1]
ndims
```

We can visualize this on the elbow plot :

```{r elbowplot, fig.width = 12, fig.height = 4}
elbow_p = Seurat::ElbowPlot(sobj, ndims = 100, reduction = "RNA_pca") +
  ggplot2::geom_point(x = ndims, y = stdev[ndims], col = "red")
x_text = ggplot_build(elbow_p)$layout$panel_params[[1]]$x$get_labels() %>% as.numeric()
elbow_p = elbow_p +
  ggplot2::scale_x_continuous(breaks = sort(c(x_text, ndims)), limits = c(0, 100))
x_color = ifelse(ggplot_build(elbow_p)$layout$panel_params[[1]]$x$get_labels() %>%
                   as.numeric() %>% round(., 2) == round(ndims, 2), "red", "black")
elbow_p = elbow_p +
  ggplot2::theme_classic() +
  ggplot2::theme(axis.text.x = element_text(color = x_color))

elbow_p
```

We generate a tSNE and a UMAP with `r ndims` principal components :

```{r tsne_umap, time_it = TRUE}
sobj = Seurat::RunTSNE(sobj,
                       reduction = "RNA_pca",
                       dims = 1:ndims,
                       seed.use = 1337L,
                       num_threads = n_threads, # Rtsne::Rtsne option
                       reduction.name = paste0("RNA_pca_", ndims, "_tsne"))

sobj = Seurat::RunUMAP(sobj,
                       reduction = "RNA_pca",
                       dims = 1:ndims,
                       seed.use = 1337L,
                       reduction.name = paste0("RNA_pca_", ndims, "_umap"))
```

We can visualize the two representations :

```{r see_umap_tsne, fig.width = 8, fig.height = 4, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "project_name",
                       reduction = paste0("RNA_pca_", ndims, "_tsne")) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("PCA - tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "project_name",
                       reduction = paste0("RNA_pca_", ndims, "_umap")) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("PCA - UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```

## Batch-effect correction

We remove sample specific effect on the pca using Harmony :

```{r harmony, fig.width = 8, fig.height = 5, time_it = TRUE}
`%||%` = function(lhs, rhs) {
  if (!is.null(x = lhs)) {
    return(lhs)
  } else {
    return(rhs)
  }
}

set.seed(1337L)
sobj = harmony::RunHarmony(object = sobj,
                           group.by.vars = "project_name",
                           plot_convergence = TRUE,
                           reduction = "RNA_pca",
                           assay.use = "RNA",
                           reduction.save = "harmony",
                           max.iter.harmony = 50,
                           project.dim = FALSE)
```

From this batch-effect removed projection, we generate a tSNE and a UMAP.

```{r harmony_tsne_umap, time_it = TRUE}
sobj = Seurat::RunUMAP(sobj, 
                       seed.use = 1337L,
                       dims = 1:ndims,
                       reduction = "harmony",
                       reduction.name = paste0("harmony_", ndims, "_umap"),
                       reduction.key = paste0("harmony_", ndims, "umap_"))

sobj = Seurat::RunTSNE(sobj,
                       dims = 1:ndims,
                       seed.use = 1337L,
                       num_threads = n_threads, # Rtsne::Rtsne option
                       reduction = "harmony",
                       reduction.name = paste0("harmony_", ndims, "_tsne"),
                       reduction.key = paste0("harmony", ndims, "tsne_"))
```

We visualize the corrected projections :

```{r see_umap_tsne_after, fig.width = 8, fig.height = 4, class.source = "fold-hide"}
tsne = Seurat::DimPlot(sobj, group.by = "project_name",
                       reduction = paste0("harmony_", ndims, "_tsne")) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("PCA - harmony - tSNE") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5),
                 legend.position = "none")

umap = Seurat::DimPlot(sobj, group.by = "project_name",
                       reduction = paste0("harmony_", ndims, "_umap")) +
  ggplot2::scale_color_manual(values = sample_info$color,
                              breaks = sample_info$project_name) +
  Seurat::NoAxes() + ggplot2::ggtitle("PCA - harmony - UMAP") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

tsne | umap
```

We will keep the tSNE from harmony :

```{r set_name2D}
reduction = "harmony"
name2D = paste0("harmony_", ndims, "_tsne")
```


## Clustering

We generate a clustering :

```{r clustering, fig.width = 6, fig.height = 6}
sobj = Seurat::FindNeighbors(sobj, reduction = reduction, dims = 1:ndims)
sobj = Seurat::FindClusters(sobj, resolution = 1.5)

clusters_plot = Seurat::DimPlot(sobj, reduction = name2D, label = TRUE) +
  Seurat::NoAxes() + Seurat::NoLegend() +
  ggplot2::labs(title = "Clusters ID") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))
clusters_plot
```


# Visualization

We represent the 4 quality metrics :

```{r qc_plot, fig.width = 12, fig.height = 3}
plot_list = Seurat::FeaturePlot(sobj, reduction = name2D,
                                combine = FALSE, pt.size = 0.25,
                                features = c("percent.mt", "percent.rb", "log_nCount_RNA", "nFeature_RNA"))
plot_list = lapply(plot_list, FUN = function(one_plot) {
  one_plot +
    Seurat::NoAxes() +
    ggplot2::scale_color_gradientn(colors = aquarius:::color_gene) +
    ggplot2::theme(aspect.ratio = 1)
})

patchwork::wrap_plots(plot_list, nrow = 1)
```


## Clusters

We can represent clusters, split by sample of origin :

```{r plot_split_dimred_cluster, fig.width = 14, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "project_name",
                                        group_by = "seurat_clusters",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = aquarius::gg_color_hue(length(levels(sobj$seurat_clusters))))

plot_list[[length(plot_list) + 1]] = clusters_plot +
  ggplot2::labs(title = "Cluster ID") &
  ggplot2::theme(plot.title = element_text(hjust = 0.5, size = 15))

patchwork::wrap_plots(plot_list, ncol = 4) &
  Seurat::NoLegend()
```

We make a heatmap to compare the representativness of cells for each sample, within each cluster :

```{r cluster_by_sample_heatmap, fig.width = 10, fig.height = 5, class.source = "fold-hide"}
cluster_by_sample = table(sobj$sample_identifier,
                          sobj@meta.data[, "seurat_clusters"]) %>%
  prop.table(margin = 1) %>%
  as.matrix()

## Representative markers
cluster_markers = c("PTPRC", "CD3E", "AIF1", "MSX2", "DIO2", "GPX2", "KRT16")
ht_annot = Seurat::GetAssayData(sobj, assay = "RNA", slot = "data")[cluster_markers, ] %>%
  Matrix::t() %>% as.data.frame()
ht_annot$clusters = sobj@meta.data[, "seurat_clusters"]
ht_annot = ht_annot %>%
  dplyr::group_by(clusters) %>%
  dplyr::summarise_all(funs('mean' = mean)) %>%
  as.data.frame() %>%
  dplyr::select(-clusters) %>%
  `colnames<-`(c(cluster_markers))

color_fun = function(one_gene) {
  gene_range = range(ht_annot[, one_gene])
  gene_palette = circlize::colorRamp2(colors = c("#FFFFFF", aquarius::color_gene[-1]),
                                      breaks = seq(from = gene_range[1], to = gene_range[2],
                                                   length.out = length(aquarius::color_gene)))
  return(gene_palette)
}

## Bottom annotation : average gene expression
ha_bottom = ComplexHeatmap::HeatmapAnnotation(df = ht_annot,
                                              which = "column",
                                              show_legend = FALSE,
                                              col = setNames(nm = cluster_markers,
                                                             lapply(cluster_markers, FUN = color_fun)),
                                              annotation_name_side = "left")

## Right annotation : number of cells by dataset
ht_annot = table(sobj$sample_identifier) %>%
  as.data.frame.table() %>%
  `colnames<-`(c("sample_identifier", "nb_cells")) %>%
  `rownames<-`(.$sample_identifier) %>%
  dplyr::select(-sample_identifier)

ha_right = ComplexHeatmap::HeatmapAnnotation(
  df = ht_annot,
  which = "row",
  show_legend = TRUE,
  annotation_name_side = "top",
  col = list(nb_cells  = circlize::colorRamp2(colors = RColorBrewer::brewer.pal(name = "Greys", n = 9),
                                              breaks = seq(from = range(ht_annot$nb_cells)[1],
                                                           to = range(ht_annot$nb_cells)[2],
                                                           length.out = 9))))

## Top annotation : main cell type in this cluster
ht_annot = table(sobj$sample_identifier,
                 sobj@meta.data[, "seurat_clusters"]) %>%
  prop.table(margin = 1) %>%
  as.matrix()

ht_annot = table(sobj$cell_type,
                 sobj$seurat_clusters) %>%
  prop.table(., margin = 2) %>%
  apply(., 2, which.max)
ht_annot = data.frame(row.names = names(ht_annot),
                      cell_type = names(color_markers)[ht_annot])

ha_top = ComplexHeatmap::HeatmapAnnotation(
  df = ht_annot,
  which = "column",
  show_legend = TRUE,
  annotation_name_side = "left",
  col = list(cell_type = color_markers))

## Assemble heatmap
ht = ComplexHeatmap::Heatmap(cluster_by_sample,
                             heatmap_legend_param = list(title = "Proportion",
                                                         col = c("#2166AC", "#F7F7F7", "#B2182B")),
                             bottom_annotation = ha_bottom,
                             right_annotation = ha_right,
                             top_annotation = ha_top,
                             cluster_rows = TRUE,
                             cluster_columns = TRUE,
                             row_title = "Sample",
                             row_names_gp = grid::gpar(names = sample_info$sample_identifier,
                                                       col = sample_info$color,
                                                       fontface = "bold"),
                             column_title = "Cluster",
                             row_names_side = "left",
                             column_names_side = "top",
                             column_names_rot = 0)

## Draw !
ComplexHeatmap::draw(ht, merge_legends = TRUE)
```

## Cell type

We visualize cell type :

```{r see_cell_type, fig.width = 10, fig.height = 8}
plot_list = lapply((c(paste0("RNA_pca_", ndims, "_tsne"),
                      paste0("RNA_pca_", ndims, "_umap"),
                      paste0("harmony_", ndims, "_tsne"),
                      paste0("harmony_", ndims, "_umap"))),
                   FUN = function(one_red) {
                     Seurat::DimPlot(sobj, group.by = "cell_type",
                                     reduction = one_red,
                                     cols = color_markers) +
                       Seurat::NoAxes() + ggplot2::ggtitle(one_red) +
                       ggplot2::theme(aspect.ratio = 1,
                                      plot.title = element_text(hjust = 0.5))
                   })

patchwork::wrap_plots(plot_list, nrow = 2) +
  patchwork::plot_layout(guides = "collect")
```


We make a representation split by origin to show cell types :

```{r cell_type_split, fig.width = 14, fig.height = 8}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "project_name",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_by = "cell_type",
                                        group_color = color_markers)

plot_list[[length(plot_list) + 1]] = patchwork::guide_area()

patchwork::wrap_plots(plot_list, ncol = 4) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "right")
```

## Cell cycle

We visualize cell cycle annotation, and BIRC5 and TOP2A expression levels  :

```{r cell_cycle, fig.width = 10, fig.height = 8, class.source = "fold-hide"}
plot_list = list()

# Seurat
plot_list[[1]] = Seurat::DimPlot(sobj, group.by = "Seurat.Phase",
                                 reduction = name2D) +
  Seurat::NoAxes() + ggplot2::labs(title = "Seurat annotation") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

# cyclone
plot_list[[2]] = Seurat::DimPlot(sobj, group.by = "cyclone.Phase",
                                 reduction = name2D) +
  Seurat::NoAxes() + ggplot2::labs(title = "cyclone annotation") +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

# BIRC5
plot_list[[3]] = Seurat::FeaturePlot(sobj, features = "BIRC5",
                                     reduction = name2D) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

# TOP2A
plot_list[[4]] = Seurat::FeaturePlot(sobj, features = "TOP2A",
                                     reduction = name2D) +
  ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1,
                 plot.title = element_text(hjust = 0.5))

patchwork::wrap_plots(plot_list, ncol = 2)
```

# Save

We save the Seurat object :

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

