---
title: "HS project"
subtitle: "Zoom in IFE granular and spinous"
author: "Audrey"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    code_folding: show
    code_download: true
    toc: true
    toc_float: true
    number_sections: false
---

<style>
body {
text-align: justify}
</style>

<!-- Automatically computes and prints in the output the running time for any code chunk -->
```{r, echo=FALSE}
# https://github.com/rstudio/rmarkdown/issues/1453
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold_", type)]])) return(res)
    
    paste0(
      "<details><summary>", "show", "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot"),
  time_it = local({
    now = NULL
    function(before, options) {
      if (options$time_it) {
        if (before) {
          now <= Sys.time()
        } else {
          res = difftime(Sys.time(), now, units = "secs")
          paste("(Time to run :", round(res, digits = 2), "s)")
        }
      }
    }
  })
)
```

<!-- Set default parameters for all chunks -->
```{r, setup, include = FALSE}
set.seed(1337L)
knitr::opts_chunk$set(echo = TRUE, # display code
                      # display chunk output
                      message = FALSE,
                      warning = FALSE,
                      fold_output = FALSE, # usefull for sessionInfo()
                      fold_plot = FALSE,
                      
                      # figure settings
                      fig.align = 'center',
                      fig.width = 20,
                      fig.height = 15,
                      
                      # something about seed, chunk and Rmarkdown compilation
                      # https://stackoverflow.com/questions/39417003/long-vectors-not-supported-yet-error-in-rmd-but-not-in-r-script
                      # cache = TRUE,
                      cache.lazy = FALSE, 
                      
                      # add runtime after chunk
                      time_it = FALSE)
```


This file is used to analyse the IFE granular and spinous dataset.

```{r library}
library(dplyr)
library(patchwork)
library(ggplot2)
library(ComplexHeatmap)

.libPaths()
```

# Preparation

In this section, we set the global settings of the analysis. We will store data there :

```{r out_dir}
save_name = "ifegs"
out_dir = "."
```

We load the dataset :

```{r load_sobj}
sobj = readRDS(paste0(out_dir, "/", save_name, "_sobj.rds"))
sobj
```

We load the sample information :

```{r custom_palette_sample, fig.width = 6, fig.height = 6}
sample_info = readRDS(paste0(out_dir, "/../../1_metadata/hs_hd_sample_info.rds"))
project_names_oi = sample_info$project_name

graphics::pie(rep(1, nrow(sample_info)),
              col = sample_info$color,
              labels = sample_info$project_name)
```

Here are custom colors for each cell type :

```{r color_markers, fig.width = 10, fig.height = 1.2, class.source = "fold-hide"}
color_markers = readRDS(paste0(out_dir, "/../../1_metadata/hs_hd_color_markers.rds"))

data.frame(cell_type = names(color_markers),
           color = unlist(color_markers)) %>%
  ggplot2::ggplot(., aes(x = cell_type, y = 0, fill = cell_type)) +
  ggplot2::geom_point(pch = 21, size = 5) +
  ggplot2::scale_fill_manual(values = unlist(color_markers), breaks = names(color_markers)) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position = "none",
                 axis.line = element_blank(),
                 axis.title = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text.y = element_blank(),
                 axis.text.x = element_text(angle = 30, hjust = 1))
```

This is the projection of interest :

```{r name2D}
name2D = "harmony_22_tsne"
```

We design a custom function to make the GSEA plot and a word cloud graph :

```{r make_gsea_plot, class.source = "fold-hide"}
make_gsea_plot = function(gsea_results, gs_oi, fold_change, metric = "FC") {
  fold_change$metric = fold_change[, metric]
  
  plot_list = lapply(gs_oi, FUN = function(gene_set) {
    # Gene set content
    gs_content = gene_sets %>%
      dplyr::filter(gs_name == gene_set) %>%
      dplyr::pull(ensembl_gene) %>%
      unique()
    
    # Gene set size
    nb_genes = length(gs_content)
    
    # Enrichment metrics
    NES = gsea_results@result[gene_set, "NES"]
    p.adjust = gsea_results@result[gene_set, "p.adjust"] %>%
      round(., 4)
    qvalues = gsea_results@result[gene_set, "qvalues"]
    
    if (p.adjust > 0.05) {
      p.adjust = paste0("<span style='color:red;'>", p.adjust, "</span>")
    }
    
    my_subtitle = paste0("\nNES : ", round(NES, 2),
                         " | padj : ", p.adjust,
                         " | qval : ", round(qvalues, 4),
                         " | set size : ", nb_genes, " genes")
    
    # Size limits
    lower_FC = min(fold_change[gs_content, ]$metric, na.rm = TRUE)
    upper_FC = max(fold_change[gs_content, ]$metric, na.rm = TRUE)
    
    # Plot
    p = enrichplot::gseaplot2(x = gsea_results, geneSetID = gene_set) +
      ggplot2::labs(title = gene_set,
                    subtitle = my_subtitle) +
      ggplot2::theme(plot.title = element_text(hjust = 0.5, face = "bold",
                                               margin = ggplot2::margin(3, 3, 5, 3)),
                     plot.subtitle = ggtext::element_markdown(hjust = 0.5,
                                                              size = 10))
    
    wc = ggplot2::ggplot(fold_change[gs_content, ],
                         aes(label = gene_name, size = abs(metric), color = metric)) +
      ggwordcloud::geom_text_wordcloud_area(show.legend = TRUE) +
      ggplot2::scale_color_gradient2(
        name = metric,
        low = aquarius::color_cnv[1],
        mid = "gray70", midpoint = 0,
        high = aquarius::color_cnv[3]) +
      ggplot2::scale_size_area(max_size = 7) +
      ggplot2::theme_minimal() +
      ggplot2::guides(size = "none")
    
    return(list(p, wc))
  }) %>% unlist(., recursive = FALSE)
  
  return(plot_list)
}
```

# Visualization

## Gene expression

We visualize gene expression for some markers :

```{r plot_list_features, fig.width = 12, fig.height = 7}
features = c("percent.mt", "nFeature_RNA", "TK1",
             "SPINK5", "KRT1", "KRTDAP")

plot_list = lapply(features, FUN = function(one_gene) {
  Seurat::FeaturePlot(sobj, features = one_gene,
                      reduction = name2D) +
    ggplot2::theme(aspect.ratio = 1) +
    ggplot2::scale_color_gradientn(colors = aquarius::color_gene) +
    Seurat::NoAxes()
})

patchwork::wrap_plots(plot_list, ncol = 3)
```

## Clusters

We visualize clusters :

```{r see_clustering, fig.width = 6, fig.height = 4}
cluster_plot = Seurat::DimPlot(sobj, reduction = name2D, label = TRUE) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)
cluster_plot
```

## Cluster annotation

KRT1 and KRTDAP being specifically expressed in IFE spinous, we make a cluster annotation:

```{r cluster_type}
sobj$cluster_type = case_when(
  sobj$seurat_clusters %in% c(5,6) ~ "IFE proliferative",
  sobj$seurat_clusters %in% c(1,2,3) ~ "IFE spinous",
  sobj$seurat_clusters %in% c(0,4) ~ "IFE granular",
  TRUE ~ "others") %>% # others: 8,9,7,0
  factor(., levels = c("IFE spinous", "IFE granular",
                       "IFE proliferative", "others"))

cluster_color = setNames(nm = c("IFE spinous", "IFE granular", "IFE proliferative", "others"),
                         c("steelblue4", "lightskyblue", "gray50", "gray10"))
```

We visualize the cluster annotation split by sample :

```{r plot_split_dimred, fig.width = 12, fig.height = 7}
plot_list = aquarius::plot_split_dimred(sobj,
                                        reduction = name2D,
                                        split_by = "project_name",
                                        group_by = "cluster_type",
                                        split_color = setNames(sample_info$color,
                                                               nm = sample_info$project_name),
                                        group_color = cluster_color,
                                        bg_pt_size = 0.5, main_pt_size = 0.5)

plot_list[[length(plot_list) + 1]] = Seurat::DimPlot(sobj, reduction = name2D,
                                                     group.by = "cluster_type") +
  ggplot2::scale_color_manual(values = cluster_color,
                              breaks = names(cluster_color)) +
  Seurat::NoAxes() +
  ggplot2::theme(aspect.ratio = 1)

patchwork::wrap_plots(plot_list, ncol = 4) &
  Seurat::NoLegend()
```

# Differential expression

We save the results in a list :

```{r list_results}
list_results = list()
```

We make over-representation analysis for each group of genes. We load gene sets from MSigDB :

```{r gene_sets}
gene_sets = aquarius::get_gene_sets(species = "Homo sapiens")
gene_sets = gene_sets$gene_sets

head(gene_sets)
```

How many gene sets ?

```{r count_gene_sets}
gene_sets[, c("gs_subcat", "gs_name")] %>%
  dplyr::distinct() %>%
  dplyr::pull(gs_subcat) %>%
  table() %>%
  as.data.frame.table() %>%
  `colnames<-`(c("Category", "Nb gene sets"))
```

We get gene name and gene ID correspondence :

```{r gene_corresp}
gene_corresp = sobj@assays[["RNA"]]@meta.features[, c("gene_name", "Ensembl_ID")] %>%
  `colnames<-`(c("NAME", "ID")) %>%
  dplyr::mutate(ID = as.character(ID))
rownames(gene_corresp) = gene_corresp$ID

head(gene_corresp)
```

## Granular vs spinous

We perform a differential expression analysis between granular and spinous clusters.

```{r set_idents1}
group_name = "granular_vs_spinous"

Seurat::Idents(sobj) = sobj$cluster_type
table(Seurat::Idents(sobj))
```

We visualize these populations :

```{r see_red_blue, fig.width = 7, fig.height = 7}
aquarius::plot_red_and_blue(sobj, reduction = name2D,
                            group1 = "IFE granular", group2 = "IFE spinous")
```

### DE

We identify DE genes between IFE granular and IFE spinous :

```{r de_ifegs, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj, ident.1 = "IFE granular", ident.2 = "IFE spinous")

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[[group_name]]$mark = mark

dim(mark)
head(mark, n = 20)
```


### Up in IFE granular

We explore enrichment in gene sets for HS population.

```{r ora_ifegs_hs, fig.width = 12, fig.height = 6}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = "Up-regulated in IFE granular compared to IFE spinous")

list_results[[group_name]]$enrichr_hs = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

### Up in IFE spinous

We explore enrichment in gene sets for HD population.

```{r ora_ifegs_hd, fig.width = 12, fig.height = 6}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = "Down-regulated in IFE granular compared to IFE spinous")

list_results[[group_name]]$enrichr_hd = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


### GSEA

We run a GSEA for all gene sets, from the full count matrix :

```{r gsea_in_ifegs, fig.width = 12, fig.height = 40}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% "IFE granular"],
                                            group2 = colnames(sobj)[sobj@active.ident %in% "IFE spinous"])
names(ranked_gene_list) = gene_corresp$ID

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(., n = 200, wt = abs(NES)) %>%
  dplyr::mutate(too_long = ifelse(nchar(ID) > 60, yes = TRUE, no = FALSE)) %>%
  dplyr::mutate(ID = stringr::str_sub(ID, end = 60)) %>%
  dplyr::mutate(ID = ifelse(too_long, yes = paste0(ID, "..."), no = ID)) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

We compute a fold change table to make a wordcloud :

```{r fold_change_ifegs}
fold_change = gene_corresp
colnames(fold_change) = c("gene_name", "ID")
rownames(fold_change) = fold_change$ID
fold_change$FC = ranked_gene_list[rownames(fold_change)]
fold_change$pct.1 = Seurat::GetAssayData(sobj, assay = "RNA",
                                         slot = "counts")[, sobj@active.ident == "IFE granular"] %>%
  apply(., MARGIN = 1, FUN = function(one_row) {
    return( mean(one_row != 0) )
  })
fold_change$pct.2 = Seurat::GetAssayData(sobj, assay = "RNA",
                                         slot = "counts")[, sobj@active.ident == "IFE spinous"] %>%
  apply(., MARGIN = 1, FUN = function(one_row) {
    return( mean(one_row != 0) )
  })
fold_change$FC_x_pct = ifelse(fold_change$FC > 0,
                              yes = fold_change$FC * fold_change$pct.1,
                              no = fold_change$FC * fold_change$pct.2)

dim(fold_change) ; head(fold_change)
```

We make the gsea plot for some gene sets :

```{r gsea_plot_within_ifegs, fig.width = 12, fig.height = 25}
gs_oi = c("GOBP_KERATINIZATION",
          "GOCC_CORNIFIED_ENVELOPE",
          "GOBP_CELLULAR_RESPONSE_TO_COPPER_ION",
          "GOBP_CELLULAR_RESPONSE_TO_CADMIUM_ION",
          "REACTOME_RESPONSE_TO_METAL_IONS",
          "WP_TGFBETA_RECEPTOR_SIGNALING")

plot_list = make_gsea_plot(gsea_results, gs_oi, fold_change, "FC_x_pct")

patchwork::wrap_plots(plot_list, ncol = 2, widths = c(2,1.5))
```

We attribute score for each gene sets to all cells, to visualize the scores as violin plots.

```{r score_gene_sets}
for (one_gene_set in gs_oi) {
  new_column = paste0("score_", one_gene_set)
  
  # gene set content
  gs_content = gene_sets %>%
    dplyr::filter(gs_name == one_gene_set) %>%
    dplyr::pull(gene_symbol) %>%
    unlist()
  
  # add score
  sobj@meta.data[, new_column] = Seurat::AddModuleScore(sobj,
                                                        features = list(gs_content))$Cluster1
}

head(sobj@meta.data)
```

We visualize the scores as violin plots :

```{r see_score_gsea, fig.width = 12, fig.height = 12}
plot_list = lapply(gs_oi, FUN = function(one_gs) {
  p = Seurat::VlnPlot(sobj, features = paste0("score_", one_gs),
                      group.by = "cluster_type",
                      split.by = "sample_type") +
    ggplot2::scale_fill_manual(values = c("#C55F40", "#2C78E6"),
                               breaks = c("HS", "HD")) +
    ggplot2::labs(title = one_gs) +
    ggplot2::theme(axis.title.x = element_blank())
  
  return(p)
})

patchwork::wrap_plots(plot_list, ncol = 2)
```

## HS vs HD

We perform a differential expression analysis for all clusters, between HS patients (HS) and healthy donors (HD).

```{r set_idents}
group_name = "HS_vs_HD"

Seurat::Idents(sobj) = sobj$sample_type
table(Seurat::Idents(sobj))
```

### DE

We identify DE genes between HS and HD :

```{r de_ifegs_hs_vs_hd, fig.width = 5, fig.height = 5}
mark = Seurat::FindMarkers(sobj, ident.1 = "HS", ident.2 = "HD")

mark = mark %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::arrange(-avg_logFC, pct.1 - pct.2)

list_results[[group_name]]$mark = mark

dim(mark)
head(mark, n = 20)
```


### Up in HS

We explore enrichment in gene sets for HS population.

```{r ora_ifegs_hs_hs_vs_hd, fig.width = 12, fig.height = 6}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC > 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = "Up-regulated in HS compared to HD")

list_results[[group_name]]$enrichr_hs = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```

### Up in HD

We explore enrichment in gene sets for HD population.

```{r ora_ifegs_hd_hs_vs_hd, fig.width = 12, fig.height = 6}
genes_of_interest = mark %>%
  dplyr::filter(avg_logFC < 0) %>%
  rownames()

enrichr_results = aquarius::run_enrichr(gene_names = genes_of_interest,
                                        gene_corresp = gene_corresp,
                                        gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                        make_plot = TRUE,
                                        plot_title = "Down-regulated in HS compared to HD")

list_results[[group_name]]$enrichr_hd = enrichr_results$ego

enrichr_results$plot +
  ggplot2::theme(axis.text.y = element_text(size = 8))
```


### GSEA

We run a GSEA for all gene sets, from the full count matrix :

```{r gsea_in_ifegs_hs_vs_hd, fig.width = 12, fig.height = 40}
ranked_gene_list = aquarius::run_foldchange(Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts"),
                                            group1 = colnames(sobj)[sobj@active.ident %in% "HS"],
                                            group2 = colnames(sobj)[sobj@active.ident %in% "HD"])
names(ranked_gene_list) = gene_corresp$ID

gsea_results = aquarius::gsea_run(ranked_gene_list = ranked_gene_list,
                                  gene_sets = gene_sets[, c("gs_name", "ensembl_gene")],
                                  GSEA_p_val_thresh = 1)

list_results[[group_name]]$gsea = gsea_results

gsea_results@result %>%
  dplyr::filter(pvalue < 0.05) %>%
  dplyr::top_n(., n = 200, wt = abs(NES)) %>%
  dplyr::mutate(too_long = ifelse(nchar(ID) > 60, yes = TRUE, no = FALSE)) %>%
  dplyr::mutate(ID = stringr::str_sub(ID, end = 60)) %>%
  dplyr::mutate(ID = ifelse(too_long, yes = paste0(ID, "..."), no = ID)) %>%
  aquarius::gsea_plot(show_legend = TRUE) +
  ggplot2::labs(title = "GSEA using all genes (count matrix)") +
  ggplot2::theme(plot.title = element_text(size = 20))
```

We compute a fold change table to make a wordcloud :

```{r fold_change_ifegs_hs_vs_hd}
fold_change = gene_corresp
colnames(fold_change) = c("gene_name", "ID")
rownames(fold_change) = fold_change$ID
fold_change$FC = ranked_gene_list[rownames(fold_change)]
fold_change$pct.1 = Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts")[, sobj@active.ident == "HS"] %>%
  apply(., MARGIN = 1, FUN = function(one_row) {
    return( mean(one_row != 0) )
  })
fold_change$pct.2 = Seurat::GetAssayData(sobj, assay = "RNA", slot = "counts")[, sobj@active.ident == "HD"] %>%
  apply(., MARGIN = 1, FUN = function(one_row) {
    return( mean(one_row != 0) )
  })
fold_change$FC_x_pct = ifelse(fold_change$FC > 0,
                              yes = fold_change$FC * fold_change$pct.1,
                              no = fold_change$FC * fold_change$pct.2)

dim(fold_change) ; head(fold_change)
```

We make the gsea plot for some gene sets :

```{r gsea_plot_within_ifegs_hs_vs_hd, fig.width = 12, fig.height = 25}
gs_oi = c("HALLMARK_INTERFERON_ALPHA_RESPONSE",
          "HALLMARK_INTERFERON_GAMMA_RESPONSE",
          "REACTOME_KERATINIZATION",
          "GOBP_KERATINIZATION",
          "GOBP_MACROPHAGE_CYTOKINE_PRODUCTION",
          "GOCC_MHC_PROTEIN_COMPLEX",
          "KEGG_ANTIGEN_PROCESSING_AND_PRESENTATION")

plot_list = make_gsea_plot(gsea_results, gs_oi, fold_change, "FC_x_pct")

patchwork::wrap_plots(plot_list, ncol = 2, widths = c(2,1.5))
```

We attribute score for each gene sets to all cells, to visualize the scores as violin plots.

```{r score_gene_sets2}
for (one_gene_set in gs_oi) {
  new_column = paste0("score_", one_gene_set)
  
  # gene set content
  gs_content = gene_sets %>%
    dplyr::filter(gs_name == one_gene_set) %>%
    dplyr::pull(gene_symbol) %>%
    unlist()
  
  # add score
  sobj@meta.data[, new_column] = Seurat::AddModuleScore(sobj,
                                                        features = list(gs_content))$Cluster1
}

head(sobj@meta.data)
```

We visualize the scores as violin plots :

```{r see_score_gsea2, fig.width = 12, fig.height = 12}
plot_list = lapply(gs_oi, FUN = function(one_gs) {
  p = Seurat::VlnPlot(sobj, features = paste0("score_", one_gene_set),
                      group.by = "cluster_type",
                      split.by = "sample_type") +
    ggplot2::scale_fill_manual(values = c("#C55F40", "#2C78E6"),
                               breaks = c("HS", "HD")) +
    ggplot2::labs(title = one_gs) +
    ggplot2::theme(axis.title.x = element_blank())
  
  return(p)
})

patchwork::wrap_plots(plot_list, ncol = 2)
```

# Visualisation

We represent differentially expressed genes. First, we extract some DE genes :

```{r prep_genes}
features_oi = lapply(list_results, `[[`, "mark") %>%
  do.call(rbind.data.frame, .) %>%
  dplyr::filter(p_val_adj < 0.05) %>%
  dplyr::filter(pct.1 - pct.2 > 0.2 | abs(avg_logFC) > 1) %>%
  rownames() %>%
  stringr::str_split(., pattern = "\\.") %>%
  lapply(., `[[`, 2) %>%
  unlist() %>%
  unique()

length(features_oi)
```

We prepare the scaled expression matrix :

```{r mat}
mat_expression = Seurat::GetAssayData(sobj, assay = "RNA", slot = "data")[features_oi, ]
mat_expression = Matrix::t(mat_expression)
mat_expression = dynutils::scale_quantile(mat_expression) # between 0 and 1
mat_expression = Matrix::t(mat_expression)
mat_expression = as.matrix(mat_expression) # not sparse

dim(mat_expression)
```

We prepare the heatmap annotation :

```{r ha_top}
ha_top = ComplexHeatmap::HeatmapAnnotation(
  sample_type = sobj$sample_type,
  cluster_type = sobj$cluster_type,
  cluster = sobj$seurat_clusters,
  col = list(sample_type = setNames(nm = c("HS", "HD"),
                                    c("#C55F40", "#2C78E6")),
             cluster_type = cluster_color,
             cluster = setNames(nm = levels(sobj$seurat_clusters),
                                aquarius::gg_color_hue(length(levels(sobj$seurat_clusters))))))
```

And the heatmap :

```{r heatmap, fig.width = 15, fig.height = 20}
ht = ComplexHeatmap::Heatmap(mat_expression,
                             col = aquarius::color_cnv,
                             # Annotation
                             top_annotation = ha_top,
                             # Grouping
                             column_order = sobj@meta.data %>%
                               dplyr::arrange(cluster_type, sample_type, seurat_clusters) %>%
                               rownames(),
                             column_title = NULL,
                             cluster_rows = FALSE,
                             cluster_columns = FALSE,
                             show_column_names = FALSE,
                             # Visual aspect
                             show_heatmap_legend = TRUE,
                             border = TRUE)

ComplexHeatmap::draw(ht,
                     merge_legend = TRUE,
                     heatmap_legend_side = "bottom",
                     annotation_legend_side = "bottom")
```

# Save

We save the list of results :

```{r save_list_results}
saveRDS(list_results, file = paste0(out_dir, "/", save_name, "_list_results.rds"))
```

We save the Seurat object with the new annotation :

```{r save_sobj}
saveRDS(sobj, file = paste0(out_dir, "/", save_name, "_sobj_annotated.rds"))
```


# R Session

```{r sessioninfo, echo = FALSE, fold_output = TRUE}
sessionInfo()
```

